### 2014-06-23 tex/author-title-etc.tex
minor edits

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{ ERIC C.~ANDERSON$^{*,\dagger,\S}$ and PATRICK BARRY$^\ddag$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{$^*$Fisheries Ecology Division, 
    Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,
    110 Shaffer Road,
    Santa Cruz, CA 95060, USA, $^\dagger$Department of Applied Math and Statistics, University of
California, Santa Cruz, CA, $^\ddag$University of Alaska, Fairbanks, Fisheries Academic Program,
17101 Point Lena Loop Rd.,
Juneau, AK 99801, USA }

@@

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson {\em et al.}}

>>>>

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{ ERIC C.~ANDERSON$^{*,\dagger,\S}$ and PATRICK D. BARRY$^\ddag$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{$^*$Fisheries Ecology Division, 
    Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,
    110 Shaffer Road,
    Santa Cruz, CA 95060, USA, $^\dagger$Department of Applied Math and Statistics, University of
California, Santa Cruz, CA, $^\ddag$University of Alaska, Fairbanks, School of Fisheries and Ocean Science,
17101 Point Lena Loop Rd.,
Juneau, AK 99801, USA }

@@

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson \& Barry}

### 2014-06-23 tex/main-body-text.tex
minor edits

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must first specify the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value of membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is reported for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs ended up at these different solutions. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$ \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $k$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. {\sc flock} 
required 1133 minutes to complete all the runs, averaging 188.8 minutes for 
each of six groups of runs over $K$ from 2 to 6. {\sc structure} completed all the runs in 
486 minutes with each of the six groups of runs averaging 80.8 minutes. Increasing
$K$ from of 2 to 6 increased {\sc flock}'s run time by a factor of 2.4, but {\sc structure}'s
run time was increased by a factor of only 1.6.  
\hrule

The allocation of indivividuals into K clusters was congruent for each program. For all six groups 
of runs from \textit{k}=2-4, {\sc flock} lead to nearly identical clustering of individuals 
(See supplementary materials). It should be noted
that of the 50 starting allocations of individuals in the {\sc flock} procedure likelihoods of each individual in 
each reference is only reported for the 
 'best run'. Of these 'best runs' for a \textit{k} of 2 only two  
ended with identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As \textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.

Despite the similar allocation of individuals into \textit{k} clusters, the 
plateau analysis in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a \textit{k} $\geq$2, the plateau record varied greatly among each
 group for \textit{k}=2 (plateau  sequences- Run1= 3,2,3,7,4; Run2=2,5,2,3,2,3; Run3=2,2,2,2,2,5; 
Run4=2,3,4,2,4,3; Run5=2,2,5,5,3; and; Run6=2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to 
either an estimate of K$\geq$2 (Run1) or \textit{undecided} as a consequence of no plateau 
lengths being greater than 6. Interpreting the results from {\sc structure} is often accomplished by 
looking at the estimated log likelihood of the data given K, the delta k method \citep{Evannoetal2005}, and plots of 
individual \textit{$q_i$} values. Both the estimated likelihood of the data
and the plots of individuals \textit{$q_i$} values suggest that 
the most likely \textit{k} for the dataset is 6.

>>>>

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must first specify the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value of membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is reported for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs ended up at these different solutions. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$ \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed, $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. {\sc flock} 
required 1133 minutes to complete all the runs, averaging 188.8 minutes for 
each of six groups of runs over $K$ from 2 to 6. {\sc structure} completed all the runs in 
486 minutes with each of the six groups of runs averaging 80.8 minutes. Increasing
$K$ from of 2 to 6 increased {\sc flock}'s run time by a factor of 2.4, while the run time 
for  {\sc structure} increased by a factor of only 1.6.  
\hrule

The allocation of indivividuals into K clusters was congruent for each program. For all six groups 
of runs from \textit{k}=2-4, {\sc flock} lead to nearly identical clustering of individuals 
(See supplementary materials). It should be noted
that of the 50 starting allocations of individuals, the likelihood of membership for every individual 
to each cluster is only reported for the 'best run'. Of these 'best runs' for a \textit{k} of 2 only two  
ended with identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As \textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.

Despite the similar allocation of individuals into \textit{k} clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a \textit{k} $\geq$2, the plateau record varied greatly among each
 group for \textit{k}=2 (plateau  sequences- Run1= 3,2,3,7,4; Run2=2,5,2,3,2,3; Run3=2,2,2,2,2,5; 
Run4=2,3,4,2,4,3; Run5=2,2,5,5,3; and; Run6=2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to 
either an estimate of K$\geq$2 (Run1) or \textit{undecided} as a consequence of no plateau 
lengths being greater than 6. Interpreting the results from {\sc structure} is often accomplished by 
looking at the estimated log likelihood of the data given K, the delta k method \citep{Evannoetal2005}, and plots of 
individual \textit{$q_i$} values. Both the estimated likelihood of the data
and the plots of individuals \textit{$q_i$} values suggest that 
the most likely \textit{k} for the dataset is 6.
